\large **ACRONYM**: A new global measure of environmental unpredictability

### 2025-04-04 writing/appendix-1.Rmd
dropped dist figure, added color palettes fig

---
title: \large A new global measure of environmental unpredictability 
subtitle: Appendix 1
author:
  - name: Stefano Mezzini
    institute: [braes, biol]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
    correspondence: "yes"
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'ndvi-stochasticity.bib'
csl: 'freshwater-biology.csl'
fontsize: 12pt
indent: true
header-includes:
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font={scriptsize, doublespacing}, labelfont=bf, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage[left]{lineno} # for line numbers
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    pandoc_args:
# https://github.com/pandoc/lua-filters/blob/master/author-info-blocks/README.md
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: false
    keep_tex: false # for submitting TeX file
editor_options:
  chunk_output_type: console
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%')
```

@@

\noindent The figures below illustrate some of the data used for fitting the models. All color palettes below are available via the `khroma` package [@frerebeau_khroma_2024] for `R` [@r_core_team_r_2024]. The code for generating the figures is available [on GitHub](https://github.com/QuantitativeEcologyLab/ndvi-stochasticity/blob/main/analysis/figures/input-data.R).

@@

```{r polygons, fig.cap="Map of the polygons for each ecoregion. Each polygon is colored with one of 9 colors based on the polygon names. Note that some neighboring polygons have the same color."}
include_graphics('../figures/input-data/polygons.png')
```

@@

```{r distance-coast, fig.cap="Map of the estimated distance from the nearest coastal edge."}
include_graphics('../figures/input-data/dist-m.png')
```

>>>>

---
title: \large A new global measure of environmental unpredictability 
subtitle: Appendix A
author:
  - name: Stefano Mezzini
    institute: [braes, biol]
  - name: Michael J. Noonan
    email: michael.noonan@ubc.ca
    institute: [braes, biol, cmps]
    correspondence: "yes"
institute:
  - braes: Okanagan Institute for Biodiversity, Resilience, and Ecosystem Services, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - biol: Department of Biology, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
  - cmps: Department of Computer Science, Math, Physics, and Statistics, The University of British Columbia Okanagan, Kelowna, British Columbia, Canada.
bibliography: 'ndvi-stochasticity.bib'
csl: 'freshwater-biology.csl'
fontsize: 12pt
indent: true
header-includes:
    - \renewcommand{\figurename}{Figure B\!\!} # for "Figure Bx."
    - \usepackage{setspace}\doublespacing # for double-spaced text
    - \usepackage{indentfirst} # for indenting first line of each paragraph
    - \usepackage[small]{titlesec} # for smaller font for headings
    - \usepackage{caption} # for more customization of captions
    - \captionsetup[figure]{font={scriptsize, doublespacing}, labelfont=bf, aboveskip=4pt, belowskip=-15pt}
    - \usepackage{hanging} # for hanging indents in references
    - \usepackage[left]{lineno} # for line numbers
subparagraph: true # needed for \usepackage[small]{titlesec}
output:
  bookdown::pdf_document2:
    pandoc_args:
# https://github.com/pandoc/lua-filters/blob/master/author-info-blocks/README.md
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
    toc: false
    number_sections: false
    keep_tex: false # for submitting TeX file
editor_options:
  chunk_output_type: console
---

@@

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, out.width = '100%', fig.align = 'center')
```

@@

\noindent The figures below illustrate some of the data used for fitting the models. All color palettes below are available via the `khroma` package [@frerebeau_khroma_2024] for `R` [@r_core_team_r_2024]. The code for generating the figures is available [on GitHub](https://github.com/QuantitativeEcologyLab/ndvi-stochasticity/blob/main/analysis/figures/input-data.R). The figure below illustrates each color palette for each type of vision (trichromatic, deuteranope, protanope, tritanope, and achomatic). Palette A is commonly attributed to Masataka Okabe and Kei Ito [@ichihara_color_2008], while the other palettes were created by Fabio Crameri [@crameri_geodynamic_2018; @crameri_scientific_2018].

```{r color-palettes, fig.cap="Color palettes used in the figures below."}
include_graphics('../figures/input-data/color-palettes.png')
```

@@

```{r polygons, fig.cap="Map of the polygons for each ecoregion. Each polygon is colored randomly with one of the 9 colors. Note that some neighboring polygons have the same color."}
include_graphics('../figures/input-data/polygons.png')
```

@@

```{r distance-coast, fig.cap="Map of the estimated distance from the nearest coastal edge.", eval=FALSE}
include_graphics('../figures/input-data/dist-km.png')
```

### 2025-04-04 writing/manuscript.Rmd
dropped dist figure, added color palettes fig

maps of input data

NDVI palette is colorblid-friendly

@@

* stable states
* green wave surfing
* migration routes are key habitat (Ortega et al., in prep), especially in stochastic environments
* @sawyer_migratory_2019
* https://www.sciencedirect.com/science/article/pii/S0960982220308484
* https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.15169
* industrial development affects animals' ability to respond to both predictable and unpredictable change (https://www.nature.com/articles/s41559-022-01887-9)
* effects of sudden deep snow in red desert on proghorn (https://www.sciencedirect.com/science/article/abs/pii/S0960982225002957)
* https://www.nature.com/articles/s41467-023-37750-z
* https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.4238

>>>>

## Choice of color schemes

We represent NDVI using a modified version of Fabio Crameri's divergent *bukavu* palette [@crameri_ref], which has high-contrast for deuteranope and protanope vision. For NDVI values between 0 and 1 (the right half of the palette), the colors have sufficient contrast for both tritanope and achomatic vision. Appendix A contains an approximate representation of the color palette for each vision type.

## Input data

Maps are in Aappendix A.

@@

* detecting state shifts
* green wave surfing
* migration routes are key habitat (Ortega et al., in prep), especially in stochastic environments
* @sawyer_migratory_2019
* https://www.sciencedirect.com/science/article/pii/S0960982220308484
* https://onlinelibrary.wiley.com/doi/full/10.1111/gcb.15169
* industrial development affects animals' ability to respond to both predictable and unpredictable change (https://www.nature.com/articles/s41559-022-01887-9)
* effects of sudden deep snow in red desert on proghorn (https://www.sciencedirect.com/science/article/abs/pii/S0960982225002957)
* https://www.nature.com/articles/s41467-023-37750-z
* https://esajournals.onlinelibrary.wiley.com/doi/full/10.1002/ecy.4238
* see data for snow and pronghorn paper (Aikens et al. 2025)
